Notes on generating initial models for mapping into {\tt MAESTRO}

### 2010-09-03 docs/initial_models/initial_models.tex
finished corrections of initialization procedure.  text in this section can be moved elsewhere now.

\section{Creating the Initial Data from the Model Data}
In {\tt base\_state.f90}, using the subroutine {\tt
  init\_base\_state} (which is actually a terrible name since we
are computing 1D initial data, which is not quite the same thing 
as the base satte), we set the initiail
data $\rho^{\init}, T^{\init}, p^{\init}$ and $X^{\init}$ equal to the
model data.  Then, we set $p^{\init},h^{\init} =
p,h(\rho^{\init},T^{\init},X^{\init})$ through the EOS.  Note that 
we overwrite the existing value of $p^{\init}$ but this should not change
the value since we called 
$p^{\model} = p^{\model}(\rho^{\model},T^{\model},X^{\model})$ at the end of the
initial model generation routines in 
\S~\ref{Sec:Creating the Model Data from Raw Data}.

@@

@@

Given $p^{\init}, \rho^{\init}, T^{\init},$ and $X^{\init}$, this
section describes how the base state ($\rho_0$ and $p_0$) and full
state ($\rho, h, X$, and $T$) are computed.  The base state is, general, not
simply a direct copy of the initial model data, since we require that
$\rho_0 = \overline\rho$.  Additionally, we require thebase state to
be HSE according to equation (\ref{HSE Discretization}), and that the full
state is thermodynamically consistent with $p_0$.  Overall we do:
\begin{enumerate}
\item Fill $\rho^{\init}, h^{\init}, X^{\init}$, and $T^{\init}$ onto a 
  multifab to obtain the full state $\rho, h, X$, and $T$.
\item If {\tt perturb\_model} = F, skip the remaining steps and
  instead set $p_0 = p^{\init}$ and $\rho_0 = \rho^{\init}$.
\item Perturb $T$, then compute $\rho,h = \rho,h(p^{\init},T,X)$.
\item Set $\rho_0 = \overline\rho$.
\item Compute $p_0$ using equation (\ref{HSE Discretization}), using
  the $p^{\init}$ at the top of the domain as the one required boundary 
  condition.  Note that we still integrate upwards from the bottom, but then
  offset the final pressure so the pressure at the top of the domain
  remains unchanged.
\item Compute the full state $T,h = T,h(\rho,p_0,X)$.
\item Set $(\rho h)_0 = \overline{(\rho h)}$.
\end{enumerate}
Now $\rho_0 = \overline\rho$, the base state is in HSE, and the full
state is thermodynamically consistent with $p_0$.

\subsection{Multilevel Plane-Parallel}
\begin{enumerate}
\item Fill $\rho^{\init}, h^{\init}, X^{\init}$, and $T^{\init}$ onto
  a multifab to obtain the full state $\rho, h, X$, and $T$.
\item If {\tt perturb\_model} = F, skip the next two steps and instead
  set $\rho_0 = \rho^{\init}$.
\item Perturb $T$, then compute $\rho,h = \rho,h(p^{\init},T,X)$.
\item Set $\rho_0 = \overline\rho$.
\item Compute $p_0$ using equation (\ref{HSE Discretization}), using
  the pressure at the top of the domain as the initial condition.
  Note that we still integrate upwards from the bottom, but then
  offset the final pressure so the pressure at the top of the domain
  remains unchanged.  Use the coarse-fine interface discretizations
  given in \S~\ref{Sec:Coarse-Fine HSE Discretization}.
\item Compute the full state $T,h = T,h(\rho,p_0,X)$.
\item Set $(\rho h)_0 = \overline{(\rho h)}$.
\end{enumerate}
Now $\rho_0 = \overline\rho$, the base state is in HSE, and the full
state is thermodynamically consistent with $p_0$.

\subsection{Coarse-Fine HSE Discretization}\label{Sec:Coarse-Fine HSE Discretization}
When integrating the HSE discretization upward, we must use a different
differencing procedure at coarse-fine interfaces.  Figure~\ref{fig:ctof} shows
the transition from coarse (level $l-1$) to fine (level $l$), with the zone
center indices noted.  
\begin{figure}[t]
\centering
\includegraphics[width=2.5in,angle=90]{\figpath/ctof}
\caption{\label{fig:ctof} A coarse-fine interface in the 1-d base state}
\end{figure}

>>>>

\section{Creating the Initial Data from the Model Data}
In {\tt base\_state.f90}, using the subroutine {\tt
  init\_base\_state} (which is actually a terrible name since we
are computing 1D initial data, which is not quite the same thing 
as the base satte), we set the initial
data $\rho^{\init}, T^{\init}, p^{\init}$ and $X^{\init}$ equal to the
model data.  Then, we set $p^{\init},h^{\init} =
p,h(\rho^{\init},T^{\init},X^{\init})$ through the EOS.  Note that 
we overwrite the existing value of $p^{\init}$ but this should not change
the value since we called 
$p^{\model} = p^{\model}(\rho^{\model},T^{\model},X^{\model})$ at the end of the
initial model generation routines in 
\S~\ref{Sec:Creating the Model Data from Raw Data}.

@@

We deal with the edge of the star by tracking the first coarse cell 
in which the density falls below {\tt base\_cutoff\_density}.  We note 
the radius of this cell center, and the values of $\rho$, $\rho h$, $X_k$, 
$p$, and $T$ in this cells.  Then, at every level, if current cell-center 
is above this radius, we set the state equal to this stored state.  This 
ensures a consistent treatment of the edge of the star at all levels.

@@

Given $p^{\init}, \rho^{\init}, T^{\init},$ and $X^{\init}$, this
section describes how the base state ($\rho_0$ and $p_0$) and full
state ($\rho, h, X$, and $T$) are computed.  The base state is, general, not
simply a direct copy of the initial model data, since we require that
$\rho_0 = \overline\rho$.  Additionally, we require thebase state to
be HSE according to equation (\ref{HSE Discretization}), and that the full
state is thermodynamically consistent with $p_0$.  Overall we do:
\begin{enumerate}
\item Fill $\rho^{\init}, h^{\init}, X^{\init}$, and $T^{\init}$ onto a 
  multifab to obtain the full state $\rho, h, X$, and $T$.
\item If {\tt perturb\_model} = T, a user-defined perturbation is
  added.  This routine should make sure that the EOS is called so that
  there is some sense of thermodynamic consistency.
\item Set $\rho_0 = \overline\rho$.
\item Compute $p_0$ using {\tt enforce\_HSE}.
\item Compute $T,h = T,h(\rho,p_0,X_k)$.
\item Set $(\rho h)_0 = \overline{(\rho h)}$.
\item Compute $\overline{T}$.  Note that we only use $\overline{T}$ as
  a diagnostic and as a seed for EOS calls.
\end{enumerate}
Now $\rho_0 = \overline\rho$, the base state is in HSE, and the full
state is thermodynamically consistent with $p_0$.

\subsection{Coarse-Fine {\tt enforce\_HSE} Discretization}\label{Sec:Coarse-Fine HSE Discretization}
When integrating the HSE discretization upward, we must use a different
differencing procedure at coarse-fine interfaces.  Figure~\ref{fig:ctof} shows
the transition from coarse (level $l-1$) to fine (level $l$), with the zone
center indices noted.  
\begin{figure}[t]
\centering
\includegraphics[width=2.5in,angle=90]{\figpath/ctof}
\caption{\label{fig:ctof} A coarse-fine interface in the 1-d base state}
\end{figure}
